---
title: "Promotion Sample"
author: "Rachel"
date: "2021/4/20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load packages:
```{r, include = FALSE}
library(tidyverse)
library(ggplot2)
library(reshape)
library(lattice)
library(DataExplorer)
library(corrplot)
library(Hmisc)
library(lme4)
library(nlme)
```

#On current employees:
```{r, include = FALSE}
black_saber_current_employees <- read_csv("rstudio-export/black-saber-current-employees.csv")
current_employees <- black_saber_current_employees

summary(current_employees)
```

#change data type for wrangling: 
```{r}
current_employees$employee_id <- as.character(current_employees$employee_id)
current_employees$salary <- gsub("\\$", "", current_employees$salary)
current_employees$salary <- gsub("\\,", "", current_employees$salary)
current_employees$salary <- as.numeric(current_employees$salary)
```
##Transform Data: #of promotime, avg. productivity,salary increase:

#create subset for # of promotion:
```{r}
num_promoted <- current_employees %>% select(employee_id,role_seniority)
wip1 <- num_promoted %>% 
  group_by(employee_id) %>% count(role_seniority) %>%
  pivot_wider(names_from = role_seniority, values_from = n)


wip1$na_count <- apply(wip1, 1, function(x) sum(is.na(x)))
wip1$num_promoted <- 9 - wip1$na_count
promoted <- wip1[,c(1,12)] 
wip1$num_promoted <- wip1$num_promoted -1
```

#average productivity for every employee:
```{r}
avg_prod <- current_employees %>% group_by(employee_id) %>% summarise(mean(productivity))
```

#calculate salary increase for every employee:
```{r}
wip2<- current_employees %>% group_by(employee_id) %>% summarise(max(salary))
wip3 <- current_employees %>% group_by(employee_id) %>% summarise(min(salary))
salary_inc <- inner_join(wip2,wip3,mby="employee_id")
salary_inc$raise <- salary_inc$`max(salary)`-salary_inc$`min(salary)`
```
#Re-join all new features with unique employee ID: 
```{r}
all_employee <- inner_join(avg_prod,promoted,by="employee_id")
all_employee <- inner_join(all_employee,salary_inc,by="employee_id")
all_employee <- all_employee[,c(1,2,3,6)]
all_employee$salary_raise <- all_employee$raise
all_employee$avg_productivity <- all_employee$`mean(productivity)`
all_employee <- all_employee[,c(1,3,5,6)]

other_subset <- current_employees[,c(1:3, 6)]
leadership <- other_subset %>% 
  group_by(employee_id) %>% count(leadership_for_level) %>%
  pivot_wider(names_from = leadership_for_level, values_from = n) 
leadership[is.na(leadership)] <- 0

other_subset <- other_subset[,c(1:3)]
other_subset <- distinct(other_subset)

new_dat <- inner_join(all_employee, leadership, by="employee_id")
new_dat <- inner_join(new_dat, other_subset, by="employee_id")
```
#change variable data types:
```{r}
new_dat$gender <- as.factor(new_dat$gender)
new_dat$team <- as.factor(new_dat$team)

new_dat$leadership_appropriate <- new_dat$`Appropriate for level`
new_dat$leadership_needsimprov <- new_dat$`Needs improvement`
new_dat$leadership_exceeds <- new_dat$`Exceeds expectations`
new_dat <- new_dat[,-c(5:7)]

new_dat$num_promoted <- new_dat$num_promoted - 1 #

str(new_dat)
```
#Fit model:
#fixed effects & random effects:
#salary_raise as outcome variable
```{r}
LMM1 <- lme(salary_raise ~gender+team+avg_productivity,data=new_dat,random = ~1|employee_id, method = "ML")
LMM2<-lme(salary_raise ~gender+team+avg_productivity+leadership_exceeds,data=new_dat,random = ~1|employee_id, method = "ML")
```
#number of promotions as outcome variable:
```{r}
LMM3 <- lme(num_promoted ~gender+team+avg_productivity,data=new_dat,random = ~1|employee_id, method = "ML")
LMM4 <- lme(num_promoted ~gender+team+avg_productivity+leadership_exceeds,data=new_dat,random = ~1|employee_id, method = "ML")

lmtest::lrtest(LMM3, LMM4)
```


#OLD CODES FROM LAST TIME:------------------------------
#basic plot:
```{r}
#看看不同性别的avg productivity是否相差不大
new_dat %>%
  ggplot(aes(x = gender, y = avg_productivity, color = gender)) +
  geom_boxplot() +
  scale_color_manual(name = "Gender", values = c("#FDAF01", "#01B9FD", "#7501FD")) +
  theme_minimal() +
  labs(title = "Gender vs. Avg Productivity", x = "Gender", y = "Average Productivity")
```

```{r}
#看看不同性别的number of promotions是否相差不大
rbc_fig1 <- new_dat %>% 
  ggplot(aes(x = num_promoted, fill = gender)) +
  geom_bar() +
  scale_fill_manual(name = "Gender", values = c("#FDAF01", "#01B9FD", "#7501FD")) +
  theme_minimal() +
  labs(title = "Gender vs. Number of Promotions", x = "Number of Promotions", y = "Count")
rbc_fig1
ggsave("rbc fig 1.png", rbc_fig1)
```

#EDA:
#boxplot on productivity grouped by gender:
```{r}
plot_boxplot(current_employees, by = "gender", geom_boxplot_args = list("outlier.color" = "red"))
```


#leadership for level on gender:
```{r, fig.width = 5}
conclusion_fig <- current_employees %>% 
  ggplot(aes(leadership_for_level, ..count..)) +
  geom_bar(aes(fill = gender), position = "fill") +
  scale_fill_manual(name = "Gender", values = c("#FDAF01", "#01B9FD", "#7501FD")) +
  labs(title = "Distribution of Performance and Genders", y = "Proportion", x= "Performance") +
  theme_minimal()
conclusion_fig
ggsave("Conclusion Fig.png", conclusion_fig)
```

#role senority on gender:
```{r, fig.height = 5}
RBC_fig2 <- ggplot(current_employees, aes(role_seniority, ..count..)) +
  geom_bar(aes(fill = gender), position = "fill") +
  labs(title = "Role Seniority on Gender", x = "Role Seniority", y = "Count") +
  scale_fill_manual(name = "Gender", values = c("#FDAF01", "#01B9FD", "#7501FD")) +
  theme_minimal()
RBC_fig2
ggsave("RBC fig 2.png", RBC_fig2)
```

#fit model on original dataset:
```{r}
mod_current <- gls(salary ~gender+team+role_seniority+leadership_for_level+productivity,data=current_employees,method = "ML")
summary(mod_current)

mod_lg <- glm(salary ~gender+team+role_seniority+leadership_for_level+productivity,data=current_employees)
summary(mod_lg)
```

#fixed effects & random effects:
```{r}
lmm1 <- lme(salary ~gender+team+role_seniority+leadership_for_level+productivity,data=current_employees,
            random = ~1|employee_id, method = "ML")
lmm2 <- lme(salary ~gender+team+role_seniority+leadership_for_level+productivity,data=current_employees,
            random = ~1|financial_q, method = "ML")
lmm3 <- lme(salary ~gender+team+role_seniority+leadership_for_level+productivity,data=current_employees,
            random = ~1|role_seniority, method = "ML")
lmm4 <- lme(salary ~gender+team+role_seniority+leadership_for_level+productivity,data=current_employees,
            random = ~1|leadership_for_level, method = "ML")
```

#compare between models:
```{r}
anova(mod_current, lmm1, lmm2, lmm3, lmm4)
```

# QQ plots (drawn to the same scale!)
```{r}
par(mfrow = c(1,2))
lims <- c(-3.5,3.5)
qqnorm(resid(mod_current, type = "normalized"),
       xlim = lims, ylim = lims,main = "GLM base")
abline(0,1, col = "red", lty = 2)
qqnorm(resid(lmm4, type = "normalized"),
       xlim = lims, ylim = lims, main = "lmm4")
abline(0,1, col = "red", lty = 2)

summary(lmm4)
lmm4$coefficients
```

```{r}
library(DiagrammeR)
flowchart_rbc <- grViz("digraph flowchart {
      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']

      # edge definitions with the node IDs
      tab1 -> tab2;
      tab1 -> tab3;
      tab1 -> tab4;
      tab1 -> tab5;
      tab2 -> tab6;
      tab3 -> tab6;
      tab4 -> tab6;
      tab5 -> tab6
      
      
      }

      [1]: '6906 Observations of 607 employees'
      [2]: 'Gender'
      [3]: 'Team'
      [4]: 'Average Productivity'
      [5]: 'Number of Promotions'
      [6]: 'Fit Model'
      ")
flowchart_rbc
```
